---
title: "ivol: Decomposing Implied Volatilites"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ivol: Decomposing Implied Volatilites}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The package \code{ivol} can be used upon financial data to estimate time-series of risk-neutral information embedded in option implied volatilities for underlying investment vehicles such as stocks, equity-indices or other financial assets which's stochastics can be described by Brownian motion. Generally, option implied volatilities can be decomposed over two different perspectives: 

1. over degrees of moneyness <font size="1"> (= different strike levels)</font> ; `imom`
    
2. over their time-to-maturity <font size="1"> (= term structure of volatility)</font>; `ihurst` 
  
A decomposition over moneyness allows to observe higher risk-neutral moments like investor expectations on skewness and kurtosis (see e.g. Jarrow and Rudd [1982]). In difference, an analysis of the implied volatility term structure allows to examine auto-correlation expectations, which can be interpreted as risk-neutral momentum in returns.  

## About the class `ivol`

This package is used to run regression analysis on option implied volatilities and create list objects of class `ivol`. List elements are `coefficients`, `residuals`, `method`, `fitted.values` and `call`. The class supports the methods `summary`, `plot` and `print`.

## imom

Applied on implied volatilities, the function `imom` runs a quadratic regression from Gram-Chalier expansion (see Jarrow-Rudd [1982]), to estimate risk-neutral 2$^\text{nd}$ (volatility), 3$^\text{rd}$ (skewness) and 4$^\text{th}$ (kurtosis) moments. The equation is given as:
\begin{equation}
  \frac{\sigma(d)}{\sigma_{ATM}}=\hat{\alpha} + \hat{\beta}_0 \, \frac{d}{3!} + \hat{\beta}_1 \, \frac{d^2}{4!}, \qquad d = \frac{\ln K/S}{\sigma \, \sqrt{\tau}}
\end{equation}
with $K$ as the underlying price, $S$ as the option's strike level and $\sigma$ as the respective implied volatility. Consequently, $d$ captures the degree of moneyness and all volatilities are normalized by the at-the-money[^1] value on the left-hand side of the equation. The non-normality adjusted risk-neutral volatility is then represented by the regression's coefficient $\hat{\alpha}$, the 3$^\text{rd}$ moment by $\hat{\beta}_0$ and $\hat{\beta}_1$ estimates the 4$^\text{th}$ moment.

If the input data is in form of a time-series, the function runs the regression in the same frequency as the input data and returns the respective time-series of implied moments.


[^1]: Options that are arbitrarily close to at-the-money face symmetric Put and Call prices. 

#### *Example*

```{r}
library(ivol)
data(otm)
x <- imom(otm)
summary(x)
head(x$coefficients)
```


## ihurst

Fractal option pricing (cp. Hu and Oksendal [2003]) allows to decompose an underlying's option implied volatility term structure into its fractal (un-autocorrelated) volatility $\sigma_f$ and into the implied Hurst exponent $H$, measuring persistence expectations in returns. Both components can be estimated by fitting OLS regression on the log-volatility curve:
\begin{equation}
  \ln \sigma(\tau) = \hat{\alpha} + \hat{\beta} \, \ln \tau
\end{equation}
having 
\begin{equation}
  \hat{\alpha} = \exp{\sigma_f} \quad \text{and} \quad \hat{\beta} = H-0.5
\end{equation}

with $\sigma_f$ as the fractal volatility and $H$ as the implied Hurst exponent.

#### *Example*

```{r}
library(ivol)
data(atm)
x <- ihurst(atm)
plot(x)
print(x)
```

## References
Jarrow, R. and Rudd, A. [1982], 'Approximate option valuation for arbitrary stochastic processes', *Journal of Financial Economics*, **10**(3), 347-369

Hu, Y. and Oksendal, B. [2003], 'Fractional White Noise Calculus and Applications to Finance', *Infinite Dimensional Analysis, Quantum Probability and Related Topics*, **6**(1), 1-32

