---
title: "Heteroscedastic Tobit Regression"
author: "[Achim Zeileis](https://eeecon.uibk.ac.at/~zeileis/)"
date: 2017-06-13
output: rmarkdown::html_vignette
bibliography: htobit.bib
vignette: >
  %\VignetteIndexEntry{Heteroscedastic Tobit Regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{htobit, car, crch, lmtest, memisc}
  %\VignetteKeywords{heteroscedastic tobit, regression, R}
  %\VignettePackage{htobit}
---

```{r preliminaries, echo = FALSE, message = FALSE, results = "hide"}
library("htobit")
library("car")
library("crch")
library("lmtest")
library("memisc")
```

# Introduction

The [`htobit`](https://R-Forge.R-project.org/projects/uibk-rprog-2017/) package
fits tobit regression models with conditional heteroscedasticy using maximum
likelihood estimation. The model assumes an underlying latent Gaussian variable

$$y_i^* \sim \mathcal{N}(\mu_i, \sigma_i^2)$$

which is only observed if positive and zero otherwise: $y_i = \max(0, y_i^*)$.
The latent mean $\mu_i$ and scale $\sigma_i$ (latent standard deviation)
are linked to two different linear predictors
$$
\begin{aligned}
\mu_i & = x_i^\top \beta \\
\log(\sigma_i) & = z_i^\top \gamma
\end{aligned}
$$
where the regressor vectors $x_i$ and $z_i$ can be set up without restrictions,
i.e., they can be identical, overlapping or completely different or just including an intercept, etc.


See also @crch for a more detailed introduction to this model class as well as
a better implementation in the package `crch`. The main purpose of `htobit` is
to illustrate how to create such a package _from scratch_.
  
------

# Implementation

As usual in many other regression packages for R [@R], the main model fitting function `htobit()`
uses a formula-based interface and returns an (S3) object of class `htobit`:

```
htobit(formula, data, subset, na.action,
  model = TRUE, y = TRUE, x = FALSE,
  control = htobit_control(...), ...)
```

Actually, the `formula` can be a two-part `Formula` [@Formula], specifying separate sets of regressors
$x_i$ and $z_i$ for the location and scale submodels, respectively.

The underlying workhorse function is `htobit_fit()` which has a matrix interface and returns an unclassed list.

A number of standard S3 methods are provided:

| Method          | Description                                                                          |
|:----------------|:-------------------------------------------------------------------------------------|
| `print()`       | Simple printed display with coefficients                                             |
| `summary()`     | Standard regression summary; returns `summary.htobit` object (with `print()` method) |
| `coef()`        | Extract coefficients                                                                 |
| `vcov()`        | Associated covariance matrix                                                         |
| `predict()`     | (Different types of) predictions for new data                                        |
| `fitted()`      | Fitted values for observed data                                                      |
| `residuals()`   | Extract (different types of) residuals                                               |
| `terms()`       | Extract terms                                                                        |
| `model.matrix()`| Extract model matrix (or matrices)                                                   |
| `nobs()`        | Extract number of observations                                                       |
| `logLik()`      | Extract fitted log-likelihood                                                        |
| `bread()`       | Extract bread for `sandwich` covariance                                              |
| `estfun()`      | Extract estimating functions (= gradient contributions) for `sandwich` covariances   |
| `getSummary()`  | Extract summary statistics for `mtable()`                                            |
| `Boot()`        | Bootstrap regression coefficients using `car` and `boot`                             |

Due to these methods a number of useful utilities work automatically, e.g., `AIC()`, `BIC()`,
`coeftest()` (`lmtest`), `lrtest()` (`lmtest`), `waldtest()` (`lmtest`),
`linearHypothesis()` (`car`), `mtable()` (`memisc`), etc.

------

# Illustration

To illustrate the package's use in practice, a comparison of several homoscedastic
and heteroscedastic tobit regression models is applied to
data on budget shares of alcohol and tobacco for 2724 Belgian households
[taken from @Verbeek:2004]. The homoscedastic model from @Verbeek:2004 can
be replicated by:

```{r verbeek}
data("AlcoholTobacco", package = "htobit")
library("htobit")
ma <- htobit(alcohol ~ (age + adults) * log(expenditure) + oldkids + youngkids, data = AlcoholTobacco)
summary(ma)
```

This model is now modified in two directions: First, the variables influencing the location
parameter are also employed in the scale submodel. Second, because the various coefficients
for different numbers of persons in the household do not appear to be very different,
a restricted specification for this is used. Using a Wald test for testing linear hypotheses
from `car` [@car] yields

```{r restrictions}
library("car")
linearHypothesis(ma, "oldkids = youngkids")
linearHypothesis(ma, "oldkids = adults")
```

Therefore, the following models are considered:

```{r alternative}
AlcoholTobacco$persons <- with(AlcoholTobacco, adults + oldkids + youngkids)
ma2 <- htobit(alcohol ~ (age + adults) * log(expenditure) + oldkids + youngkids |
  (age + adults) * log(expenditure) + oldkids + youngkids, data = AlcoholTobacco)
ma3 <- htobit(alcohol ~ age + log(expenditure) + persons | age + log(expenditure) + persons, data = AlcoholTobacco)
BIC(ma, ma2, ma3)
```

The BIC would choose the most parsimonious model but a likelihood ratio test would prefer
the unconstrained person coefficients:

```{r lrtest}
library("lmtest")
lrtest(ma, ma2, ma3)
```

------

# Replication

To assess the reliability of the `htobit()` implementation, it is benchmarked against the `crch()`
function of @crch, using the same restricted model as above.

```{r crch}
library("crch")
ca3 <- crch(alcohol ~ age + log(expenditure) + persons | age + log(expenditure) + persons,
  data = AlcoholTobacco, left = 0)
```

Using a model table from `memisc` [@memisc] it can be easily seen the results can be
replicated using both packages:

```{r mtable}
library("memisc")
mtable("htobit" = ma3, "crch" = ca3)
```

------

# References

