---
title: "Multivariate EWMA model"
author: "Bernhard Eder"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LFU Risk Management Utilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The following vignette provides a short tutorial on the multivariate exponential weighted moving average (EWMA) model, as it is implemented in `LFUrmutils`. The multivariate EWMA model is also implemented in the package `MTS`. In comparison to MTS, our package preserves the time dimension of the used time series and also calculates the volatility for the next trading day (on the New York Stock Exchange) at the end of the time series. In addition, our package offers various convenience functions for calculating residuals and extracting the conditional variance-covariance matrices, conditional volatilities as well as conditional correlations from the output object. In contrast to MTS, it is however not yet possible to estimate the decay parameter directly from the data.

## Data analysis

First, we need to download the stock prices of several assets, and compute the corresponding log returns. For this example we use the adjusted closing prices of Microsoft, IBM and Goldman Sachs from the first of January 2010 until the 31st December 2018. Additionally, we plot the time series of the (squared) returns, to make ourselves comfortable with the data.

```{r download, fig.show='hold'}
library(LFUrmutils)
library(tseries)
library(zoo)

p1 <- get.hist.quote("msft", "2010-01-01", "2018-12-31",
                     quote = "Adjusted", quiet = TRUE)
p2 <- get.hist.quote("ibm", "2010-01-01", "2018-12-31",
                     quote = "Adjusted", quiet = TRUE)
p3 <- get.hist.quote("gs", "2010-01-01", "2018-12-31", 
                     quote = "Adjusted", quiet = TRUE)

y1 <- diff(log(p1)) * 100
y2 <- diff(log(p2)) * 100
y3 <- diff(log(p3)) * 100

y <- merge(y1, y2, y3)
head(y)
plot(y)
plot(y^2)
```

We can clearly see that there are volatility clusters present in the individual time series. Moreover, they appear at the same time, especially at Microsoft and IBM, which belong to the same industry. In addition, we could for example compute the autocorrelation functions of the squared return series, in order to gauge volatility clusters more formally, which is left as an exercise to the reader.

## Model computation and diagnostics

Now it is time to compute the multivariate EWMA model:

$$\hat{\Sigma}_t=(1-\lambda)y'_{t-1}y_{t-1}+\lambda\hat{\Sigma}_{t-1}$$

For the decay parameter $\lambda$ a value of 0.94 is chosen, as suggested by RiskMetrics (1997). However, it would be possible at least in principle to use any value between 0 and 1, although values between 0.9 and 1 are generally recommended.

```{r ewma, fig.width=7, fig.height=7}
# Compute the EWMA model
EWMA <- MultiEWMA(y, center = TRUE)

# Compare the return series used in EWMA to the original return series
head(EWMA$Returns)
head(y)

# Compute and plot the residuals
EWMAres <- residuals(EWMA, standardize = TRUE)
plot(EWMAres)

plot(EWMA$Returns[, 1], col = "gray")
lines(1.96 * sqrt(EWMA$Variances[, 1]))
lines(- 1.96 * sqrt(EWMA$Variances[, 1]))
legend("bottomright", bty = "n", lwd = 1, col = c("gray", "black"), 
       legend = c("Returns", "+/- 1.96 * Volatility"))
```

Comparing the residuals with the original return series, we can see that the model partly captures the volatility clusters, although some significant outliers remain, which is confirmed by the second plot.

## Conditional volatilities and correlations

In this section we demonstrate how we can extract the conditional volatilities and correlations. Note that each row in the output corresponds to the specified elements of conditional volatility matrix, or the conditional correlation matrix, respectively. Due to the symmetry of the matrices, it is possible to suppress redundant elements, which is particularly useful for graphical purposes. In the case of the volatility matrix, it is also possible to exclude the elements outside the diagonals. It is also possible to exclude the diagonal elements of the correlation matrix because they are always equal to one.

```{r volacor, fig.width=7, fig.height=7}
EWMAvola <- vola(EWMA, duplicates = FALSE)
plot(EWMAvola)

EWMAccor <- ccor(EWMA, diagonal = FALSE, duplicates = FALSE)
plot(EWMAccor)
```

Moreover, we can compare the conditional correlations from the multiviarate EWMA model to the 30-day rolling correlation. We carry out this analysis as an example for the returns of Microsoft and IBM. We can see that even the simple EWMA model captures the conditional correlations quite well.

```{r rollcor, fig.width=7, fig.height=7}
cor_MSFT_IBM <- rollapplyr(y[, c(1:2)], width = 30, by.column = FALSE, 
                           FUN = function(x) cor(x[, 1], x[, 2]), fill = NA)
cor_MSFT_IBM <- zoo(cor_MSFT_IBM, index(y1))
#tail(cor_MSFT_IBM)
plot(EWMAccor[, 1])
lines(cor_MSFT_IBM, col = "green", lty = 2)
```

## Compute the conditional variance-covariance matrix

In this section we demonstrate, how one can compute the conditional variance-covariance matrix based on the conditional volatilities and the conditional correlations. To this end, we need to compute and extract the diagonal elements of the volatility matrix, and the whole correlation matrix. Next, we setup an object to store the variance-covariance matrix, and compute it thereafter. Finally, we compare the computed variance-covariance matrix with the original variance-covariance matrix. As we can see, we are able to "rebuild" the conditional variance-covariance matrix.

```{r varcov}
EWMAvola <- vola(EWMA, duplicates = TRUE, offdiagonal = FALSE)
EWMAccor <- ccor(EWMA)

TT <- dim(EWMAccor)[1]
c <- sqrt(dim(EWMAccor)[2])
VarCov <- matrix(NA, dim(EWMAccor)[1], dim(EWMAccor)[2])
# diag(as.numeric(EWMAvola[1, ]))
# as.numeric(EWMAccor[1, ])
# matrix(as.numeric(EWMAccor[1, ]), c, c,  byrow = TRUE)

for(i in 1:TT){
  VarCov[i, ] <- c(diag(as.numeric(EWMAvola[i, ])) 
                   %*% matrix(as.numeric(EWMAccor[i, ]), c, c,  byrow = TRUE) 
                   %*% diag(as.numeric(EWMAvola[i, ])))
}

VarCov <- zoo(VarCov, order.by = index(EWMAvola))

tail(VarCov[, c(1:4)])
tail(EWMA$Variances[, c(1:4)])
```
